title: "Øvelser uge 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
```

## 5.1

can be represented by a tuple (c0, c1, ..., ck), e.g. the polynomial 3x3-2x2+0x+5 can be represented by a tuple (5, 0, -2, 3). Use list comprehension, enumerate, and sum to evaluate a polynomial given by its coefficients. E.g. for x = 7 and coefficients (5, 0, -2, 3) your code should compute the value 936.

```{python}
[3*x**3-2*x**2+0*x+5 for x in [7]]
```

```{python}
def poly(x, coefs):
    return sum([c * x ** k for k, c in enumerate(coefs)])

def pol2(lst, x):
    total = []
    for power, coeff in enumerate(lst):
        total.append((x ** power) * coeff)
        return sum(total)
```

```{r}
poly(7, (5, 0, -2, 3))
```


## 5.2

Given a list of first names first = ['Donald', 'Mickey', 'Scrooge'] and last names last = ['Duck', 'Mouse', 'McDuck'], use list comprehension, zip, and sorted to generate an alphabetically sorted list of names 'lastname, firstname', i.e. generates the list ['Duck, Donald', 'McDuck, Scrooge', 'Mouse, Mickey'].

```{python}
first = ['Donald', 'Mickey', 'Scrooge']
last =  ['Duck', 'Mouse', 'McDuck']
list(sorted(zip(last, first)))
```

```{python}
def zip_names(first_name, last_name):
  return [f'{last}, {first}' for last, first in sorted(zip(last_name, first_name))]

zip_names(first, last)
```

## 5.3

Implement your own version my_zip of zip, such that my_zip([L1,L2,...,Lk]) == list(zip(L1,L2,...,Lk)).

```{python}
def my_zip_snyd(L):
    return list(zip(*L))


def my_zip_comprehension(L):
    return [tuple(lst[i] for lst in L) for i in range(min([len(lst) for lst in L]))]


def my_zip_nested(L):
    min_lenght = min([len(lst) for lst in L])
    res = []
    for i in range(min_lenght):
        tup = []
        for lst in L:
            tup.append(lst[i])
        res.append(tuple(tup))
    return res
```

## 5.4

Make a method flatten(L) that given a list of lists L = [L1,L2,...,Lk] returns one list that is the concatenation of all lists, i.e. L1+L2+···+Lk. Your implementation should use list comprehension.

```{python}
def flatten1(L):
    return [ele for lst in L for ele in lst]
```


# Lecture 6 (Dictionaries and sets)

## Exercise 6.1 (histogram)

```{python}
def histogram(L):
    dic = {}
    for ele in L:
        if ele not in dic:
            dic[ele] = 0
        dic[ele] += 1
    return [(ele, num) for ele, num in dic.items()]
```


```{python}
import collections
histogram = list(collections.Counter(['A', 'B', 'A', 'A', 'C', 'E', 'C']).items())
print(histogram)
```

```{python}
def histogram(n):
  dicto = {}
  for x in n:
    if x not in dicto:
      dicto[x] = 1
    else:
      dicto[x] += 1
  return list(dict.items(dicto))

print(histogram(['A', 'B', 'A', 'A', 'C', 'E', 'C']))
```

### 6.2

```{python}
import re

txt = open("saxo.txt").read()
txt = txt.lower()
words = re.split('[^a-z]+', txt)

# list_com
most_freq = sorted(
    [(freq, word) for word, freq in histogram(words) if len(word) >= 6], reverse=True)[:10]
```

```{python}
# simplere
most_freq = []
for word, freq in histogram(words):
    if len(word) >= 6:
        most_freq.append((freq, word))
most_freq = sorted(most_freq, reverse=True)[:10]

print('Rank Freq. Word')
print('====================')
for idx, (occ, word) in enumerate(most_freq):
    print(f'{idx + 1:4d}{occ:5d} {word}')
```


## aflevering 3

```{python}
from random import randint
```

### a)

Make a function generate_labels(n), that given an integer n returns a list of n distinct strings, e.g. 'A', 'B', ... or "L1", 'L2' ...
Example: generate_labels(5) could return ['A', 'B', 'C', 'D', 'E']


```{python}
def generate_labels(n):
  letters = [chr(x) for x in range(ord('A'), ord('A') + n)]
  return letters

print("An example of the function: ", generate_labels(8))
```

### b)

Make a function permute(L), that given a list L, returns a new list containing a random permutation of the elements in L.
Hint: Construct the new list left-to-right by randomly selecting an element not selected so far. To generate a random integer in the interval [a,b], you can you the function randint(a,b) from the module random (use from random import randint to get access to the function).
Note. Using the function shuffle from the module random to solve the question would be considered cheating.
Example: permute(['A', 'B', 'C']) could return ['B', 'C', 'A'].

```{python}
def permute(n):
  for i in reversed(range(1, len(n))): 
    j = randint(0, i)  
    n[i], n[j] = n[j], n[i]
  return n
```


```{python}
test = generate_labels(2)
test
```

```{python}
permute(generate_labels(3))
```


### c)

Make a function pairs(L), that given a list of comparable elements, returns a list of all pairs, i.e. tuples with two elements, (a, b) where a < b.
Example: pairs(['A', 'F', 'B']) should return [('A', 'F'), ('A', 'B'), ('B', 'F')].

```{python}
def pairs(n):
  par = [(x,y) for x in n for y in n if x != y and x < y]
  return par 
```

```{python}
L = ['A', 'F', 'B']
pairs(['A', 'F', 'B'])
```

### d)

Make a function canonical_triplets(A, B) that returns a list of all canonical triples where the left subtree contains a label from A and the right subtree is a pair from B.
Example: canonical_triplets(['A', 'B'], ['C', 'D', 'E']) should return:
[('A', ('C', 'D')), ('A', ('C', 'E')), ('A', ('D', 'E')), ('B', ('C', 'D')), ('B', ('C', 'E')), ('B', ('D', 'E'))


```{python}
def canonical_triplets(A, B):
  can = [(x, (y,z)) for x in A for y in B for z in B if x != y and y != z and z != x and y < z]
  return can
```

```{python}
canonical_triplets(['A', 'B'], ['C', 'D', 'E'])
```

### e) 

Make a function anchored_triplets(L, R) that returns a list of all canonical triples anchored at a node where the left subtree contains the labels in the list L and the right subtree contains the labels in the list R.

```{python}
def anchored_triplets(A,B):
  tri = [(x, (y,z)) for x in A for y in B for z in B if x != y and y != z and z != x]
  return tri
```

```{python}
tri_1 = anchored_triplets(['A', 'F', 'B'], ['D', 'C', 'E'])
print(len(tri_1))
print(tri_1)
```

```{python}
from random import shuffle
import inspect
lines = inspect.getsource(shuffle)
print(lines)
```
